---
title: "Zappos.com's Customer Transactions Exploratory Data Analysis (EDA)"
author: "Dorcas Mbaeri"
date: "March 3, 2018"
output: html_document
---


```{r loadlibrary, include=FALSE}
library(tidyverse)
```


### Structure of data

This is the introduction to my senior project at the University of Houston-Downtown. We will begin by exploring the dataset containing Zappos.com's customer transactions.

But first let's load the data.

```{r mydata}
mydata <- read_csv("C:/Users/Dorcas/OneDrive - University of Houston Downtown/SeniorProject/Senior Project/Analytics_Challenge_Data.csv", col_names = TRUE, na = "NA")
```

Now, let's take look at the structure of the data:
```{r}
summary(mydata)

dim(mydata)

str(mydata)

head(mydata)
```

Here are visual distributions of some categorical variables of __*mydata*__:

Bar chart of *site*
```{r}
ggplot(data=mydata) +
  geom_bar(mapping = aes(x=site))
```

Histogram of *new_customer* (not including blanks)
```{r}
noBlanks <- mydata %>%
  filter(new_customer >= 0)

ggplot(data = noBlanks, mapping = aes(x = new_customer)) +
  geom_histogram(binwidth = 1)
```

Line graph of _platform_ and *new_customer*
```{r}
ggplot(data = noBlanks, mapping = aes(x = new_customer, y = ..density..)) +
  geom_freqpoly(mapping = aes(color = platform), binwidth = 1)
```

Distribution of the number of orders per user/customer against the site name
```{r}
ggplot(data = noBlanks, mapping = aes(x = site, y = orders)) +
  geom_boxplot()
```

Distribution of the number of orders per user/customer against a the type of platform used
```{r}
boxplot(mydata$orders ~ mydata$platform)
```
